---
title: 'Cross Correlation Analysis: Relationship between Diesel Prices and Commercial Auto Premiums'
author: "Bob McPherson"
date: "March 1, 2017"
output: word_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

##Remove All Objects before Running the Analysis
rm(list=ls())

#install.packages("Hmisc")
#install.packages("devtools")
#install.packages("Quandl")
#install.packages("car")
#install.packages("fread")
#install.packages("curl")


##Load Packages
library(Hmisc)
library(devtools)
library(Quandl) #for data sources
library(car) #for scatterplots
library(curl) #for downloand data from a url

##Load Data
########################################################
##Independent Variable

#indep.var <- Quandl("BTS/FUEL",collapse="annual")
#str(indep.var)
#indep.var2 <- indep.var[nrow(indep.var):1,] #reverse order to ascending...oldest at beginning
#x <- indep.var2[,"Total Cost per Gallon ($)"]


##Get Cellular Phone Use in US
indep.var <- Quandl("UICT/CELL_USA", api_key="DJGcfzQc5RYP1JSycMBv", api_key="DJGcfzQc5RYP1JSycMBv", collapse="annual")
str(indep.var)
indep.var2 <- indep.var[nrow(indep.var):1,] #reverse order to ascending...oldest at beginning
str(indep.var2)
x <- indep.var2$`Mobile-cellular telephone subscriptions per 100 inhabitants`
(indep.var.max.year <- max(indep.var2$Date))
(indep.var.min.year <- min(indep.var2$Date))
(indep.var.total.years <- as.numeric((indep.var.max.year - indep.var.min.year) / 365))

##Get Producer Price Index by Commodity for Fuels and Related Products and Power: No. 2 Diesel Fuel
#indep.var <- Quandl("FRED/WPU05730302", api_key="DJGcfzQc5RYP1JSycMBv", collapse="annual") 
#str(indep.var)
#indep.var2 <- indep.var[nrow(indep.var):1,] #reverse order to ascending...oldest at beginning
#x <- indep.var2$VALUE
#(indep.var.max.year <- max(indep.var2$DATE))
#(indep.var.min.year <- min(indep.var2$DATE))
#(indep.var.total.years <- as.numeric((indep.var.max.year - indep.var.min.year) / 365))

############################################################
##Dependent Variable

##Traffic crashes, traffic crash fatalities, and alcohol-related traffic crash fatalities in the United States
#dep.var <- Quandl("NIAAA/CRASHES_ALCOHOL", api_key="DJGcfzQc5RYP1JSycMBv", collapse="annual")
#str(dep.var)
#dep.var2 <- dep.var[nrow(dep.var):1,] #reverse order to ascending...oldest at beginning
#y <- dep.var2$`Traffic Crashes`
#(dep.var.max.year <- max(dep.var2$Year))
#(dep.var.min.year <- min(dep.var2$Year))
#(dep.var.total.years <- as.numeric((dep.var.max.year - dep.var.min.year) / 365))

##Get Persons Killed in Crashes Involving a Large Truck, by Person Type and Crash Type, 1994 - 2013 - State : USA
dep.var <- Quandl("NHTSA/FTR_LTR_PK_US", api_key="DJGcfzQc5RYP1JSycMBv", collapse="annual")
#str(dep.var)
dep.var2 <- dep.var[nrow(dep.var):1,] #reverse order to ascending...oldest at beginning
#str(dep.var2)
y <- dep.var2$`Multiple--Vehicle`
(dep.var.max.year <- max(dep.var2$Year))
(dep.var.min.year <- min(dep.var2$Year))
(dep.var.total.years <- as.numeric((dep.var.max.year - dep.var.min.year) / 365))


##Get Producer Price Index by Industry: Premiums for Property and Casualty Insurance: Premiums for Commercial Auto Insurance
#dep.var <- Quandl("FRED/PCU9241269241263", api_key="DJGcfzQc5RYP1JSycMBv", collapse="annual")
#str(dep.var)
#dep.var2 <- dep.var[nrow(dep.var):1,] #reverse order to ascending...oldest at beginning
#str(dep.var)
#y <- dep.var2$VALUE
#(dep.var.max.year <- max(dep.var2$DATE))
#(dep.var.min.year <- min(dep.var2$DATE))
#(dep.var.total.years <- as.numeric((dep.var.max.year - dep.var.min.year) / 365))

indep.var.max.year
dep.var.max.year

indep.var.min.year
dep.var.min.year

#diff.total.years <- floor(abs(dep.var.total.years - indep.var.total.years))

##Load Commercial Auto Accident Data from Peer Companies
#comauto <- read.csv("ComAutoPeerCos_PaidLossSNL.csv")
#y <- comauto[,3]

##shorten the longest series to match the shortest series
min.length <- min(length(x),length(y))

x <- x[((length(x)-min.length)+1):length(x)]

y <- y[((length(y)-min.length)+1):length(y)]

length(x)
length(y)
dep.var2$DATE
indep.var2$DATE

(dep.var.dates <- dep.var2$DATE[((length(dep.var2$DATE)-min.length)+1):length(dep.var2$DATE)])
(indep.var.dates <- indep.var2$DATE[((length(indep.var2$DATE)-min.length)+1):length(indep.var2$DATE)])

##minimum number of values to match in correlations
min.vector <- if(ceiling((length(x)/2)*.334) <= 4) 4 else ceiling((length(x)/2)*.334)

##function for calculating cross correlation r-values
crosscor <- function(x,y){
  R=NULL
  i=0
  for (i in (-1*min.vector):(min.vector)) {
    R <- rbind(R,cor(x,Lag(y,shift=i),use="pairwise.complete.obs"))    
  }
  R
}

##function for running p-value tests
crosscor.test <- function(x,y){
  R=NULL
  i=0
  for (i in (-1*min.vector):(min.vector)) {
    R <- rbind(R,cor.test(x,Lag(y,shift=i),use="pairwise.complete.obs")$p.value)    
  }
  R
}

```

# Examine Lags
In this section, we apply the functions above, to examine the possible lags.  The time series is offset by one period, and the correlation is tested.  This process is repeated, until the maximum number of lags is reached.  A p-value is calculated for each correlation, to test the statistical significance.  The candidate lag with the most significant p-value is selected as the best lag to use in offsetting the time series data.  The smallest p-value corresponds to the most statistically significant value.

```{r p-values}

(R <- crosscor(x,y))
(R.test <- crosscor.test(x,y)) 

plot(R,main="Plot of Correlations")

plot(R.test,main="Plot of P-Values")

(best.pval.index <- which.min(R.test))

(best.pval <- R.test[best.pval.index])

if(best.pval <= .05){
  pval.comment <- "This p-value suggests that the best lag is statistically significant at the 95% confidence level."
} else {
  pval.comment <- "This p-value suggests that the best lag is not statistically significant at the 95% confidence level."
  }

```
# Absolute Correlation Value Comparisons
Instead of comparing the p-values, we can also select the best lag based on comparing the correlation coefficients, or r-values.  In this test, we select the largest absolute value of the correlation coefficients.  In most cases, the best lag should agree with the value that was chosen using the p-value comparison.  This exhibit is merely shown to reveal the correlation coefficients behind the p-values.

```{r crosscors}

if(abs(min(R))>=max(R)){
  (max.cor <- which.min(R))
  direction <- "negative, meaning when one trend goes up, the trend for the other variable tends to move in the opposite direction"
} else {
  (max.cor <- which.max(R))
  direction <- "positive, meaning when one trend goes up or down, the trend for the other variable tends to move in the same direction"
}

(best.lag <- ((-1*min.vector):(min.vector))[max.cor])

best.correlation <- R[best.pval.index]



```
The best lag for the data is `r best.lag` years. The correlation coefficient for the best lag is `r best.correlation`.  The sign of the correlation coefficient is `r direction`.  The p-value for the best lag is `r best.pval`.  `r pval.comment`

```{r PlotLags}

#plot line graphs with no lag
#plot(x,type="l")
#par(new=T)
#plot(y,type="l")
#axis(side=4)
#par(new=F)

##plot the variables after adjusting for the lag
plot(y,type="l",col=1,xlab="Year",ylab="Dependent Var",main="Lining Up Each Trend, after Adjusting for the Best Lag")
par(new=T)
plot(Lag(x,shift=best.lag),type="l",col="blue",xlab='',ylab='',axes=F)
axis(side=4)
mtext(side=4,line=0,"Independent Var")
legend("topleft",legend=c("Dependent Var","Independent Var"),lty=c(1,1),col=c("black","blue"))
par(new=F)
```
# Auto Regressive Tendencies and First Order Differencing
Data from a time series can often have a tendency to be correlated to itself.  A common situation that results in this phenomenon is seasonality.  This is called, auto regression.  The ACF function in R tests for auto regression.  We will use this function in this section.  

We will also detrend the data using first order differencing.  This merely refers to subtracting each precending value from the value that follows.

```{r ACF}
##Rerun with first order differencing

##run ACFs
acf(x)
acf(y)

##first order differencing
x.diff <- diff(x,1)
y.diff <- diff(y,1)

##ACF on differenced data
acf(x.diff)
acf(y.diff)
```
Next, we run p-value significance tests on the cross correlations of the differenced data.  The results are also plotted.

```{r DifferencedCrossCorrelations}

(S <- crosscor(x.diff,y.diff))
(S.test <- crosscor.test(x.diff,y.diff)) 

plot(S,main="Plot of Correlations for Differenced Data")

plot(S.test,main="Plot of P-Values for Differenced Data")

if(abs(min(S))>=max(S)){
  (max.cor <- which.min(S))
  direction <- "negative, meaning when one trend goes up, the trend for the other variable tends to move in the opposite direction"
} else {
  (max.cor <- which.max(S))
  direction <- "positive, meaning when one trend goes up or down, the trend for the other     variable tends to move in the same direction"
}

best.lag <- ((-1*min.vector):(min.vector))[max.cor] 

best.correlation <- S[max.cor]

(best.pval.index <- which.min(S.test))

(best.pval <- S.test[best.pval.index])

if(best.pval <= .05){
  pval.comment <- "This p-value suggests that the best lag is statistically significant at the 95% confidence level."
} else {
  pval.comment <- "This p-value suggests that the best lag is not statistically significant at the 95% confidence level."
  }

```
The best lag for the data is `r best.lag` years. The correlation coefficient for the best lag is `r best.correlation`.  The sign of the correlation coefficient is `r direction`.  The p-value for the best lag is `r best.pval`.  `r pval.comment`

```{r PlotsOfDifferencedVersusNonDifferenced}

##plot line graph of lagged, differenced data
plot(y.diff,type="l",col=1,xlab="Year",ylab="Dependent Var",main="Lagged, Differenced Data")
par(new=T)
plot(Lag(x.diff,shift=best.lag),type="l",col="blue",xlab='',ylab='',axes=F)
axis(side=4)
mtext(side=4,line=0,"Independent Var")
legend("topleft",legend=c("Independent Var","Dependent Var"),lty=c(1,1),col=c("black","blue"))
par(new=F)

##plot line graph of lagged, NON-differenced data; 
plot(y,type="l",col=1,xlab="Year",ylab="Dependent Var",main="Lagged, Non-Differenced Data")
par(new=T)
plot(Lag(x,shift=best.lag),type="l",col="blue",xlab='',ylab='',axes=F)
axis(side=4)
mtext(side=4,line=0,"Independent Var")
legend("topleft",legend=c("Independent Var","Dependent Var"),lty=c(1,1),col=c("black","blue"))
par(new=F)

scatterplot(x.diff,Lag(y.diff,shift=best.lag),main="Lagged, Differenced Data")

scatterplot(x,Lag(y,shift=best.lag),main="Lagged, Non-Differenced Data")
```

